### 2014-07-14 paper.tex
Cleaned up excess comments.

\documentclass[11pt, oneside]{amsart}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{color}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{graphicx}
\usepackage[latin9]{inputenc}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{subfigure}
\usepackage{psfrag}
\usepackage{tabularx}
\usepackage[hyphens]{url}
\usepackage{wrapfig}
\usepackage{framed,color}
\usepackage{fancybox}
\usepackage{verbatim}

\usepackage[bookmarks, bookmarksopen, bookmarksnumbered]{hyperref}
\usepackage[all]{hypcap}
\urlstyle{rm}

\definecolor{orange}{rgb}{1.0,0.3,0.0}
\definecolor{violet}{rgb}{0.75,0,1}
\definecolor{darkgreen}{rgb}{0,0.6,0}
\definecolor{cyan}{rgb}{0.2,0.7,0.7}
\definecolor{blueish}{rgb}{0.2,0.2,0.8}

\definecolor{shadecolor}{rgb}{0.9,0.9,0.9}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\newcommand{\todo}[1]{{\color{blue}$\blacksquare$~\textsf{[TODO: #1]}}}
\newcommand{\huttonnote}[1]{ {\textcolor{magenta}    { ***Eric:      #1 }}}

\DeclareRobustCommand{\csdms}{\textsc{csdms}}
\DeclareRobustCommand{\doi}{\textsc{doi}}
\DeclareRobustCommand{\bmi}{\textsc{bmi}}
\DeclareRobustCommand{\wmt}{\textsc{wmt}}

% Don't use tt font for urls
\urlstyle{rm}

\begin{document}

\title[]{Building Sustainable Software - The CSDMS Approach}

\author{
  Eric W. H. Hutton$^{\dag}$,
  Mark D. Piper$^{\dag}$,
  Irina Overeem$^{\dag}$
  Albert J. Kettner$^{\dag}$,
}

\thanks{{}$^{\dag}$ Community Surface Dynamics Modeling System, Boulder, CO}

\begin{abstract}

\csdms, The Community Surface Dynamics Modeling System, is an NSF funded
project whose focus is to aid a diverse community of earth and ocean system
model \emph{users} and \emph{developers} to use and create robust software
quickly.  To this end, \csdms{} develops, integrates, archives and disseminates
earth-system models and tools to an international (67 country) community
with the goal of building the set of tools necessary to model the
earth system. \csdms{} is a place modelers look to for access to hundreds open
source surface-dynamics models and tools, as well as model metadata. Such a
model repository increases model transparency and helps to eliminate the
duplication of effors by presenting the current state of modeling efforts.
To increase software sustainability, composability and interoperability \csdms{}
promotes standards that define common modeling interfaces, semantic mediation
between models, and model metadata. Through online resources and workshops,
\csdms{} promotes software engineering best practices, which are foreign to many
developers within our modeling community. For example, version control, unit
testing, continious integration, test-driven development, well-written
\emph{clean code} are all topics of the educational mission of \csdms{}.

\end{abstract}

\maketitle

\section{The Community Surface Dynamics Modeling System}

The mission of the Community Surface Dynamics Modeling System (\csdms{})
~\cite{peckham2012component} is to help a diverse modeling community toward
common goals and standards. This effort involves creating:
\begin{itemize}
  \item a repository of source code and metadata for open-source models and
    tools (Section~\ref{sec:organizing})
  \item reusable plug-and-play model components and a framework
        within which they can be coupled to create new models
        (Section~\ref{sec:reusable})
  \item an \emph{efficient} and \emph{open} modeling community through
        standards (Section ~\ref{sec:standards}) and education
        (Section~\ref{sec:education})
\end{itemize}

In building a modeling framework, \csdms{} has leveraged
several existing, well-established and open-source software tools. For example,
\csdms{} uses tools from the Common Component Architecture (CCA)
~\cite{armstrong1999toward} toolchain: \emph{Babel}, and \emph{Bocca}. Babel
provides interoperability between components
written in different languages; it currently supports C, C++, Fortran, Java,
and Python. Bocca helps with creating CCA-compliant components and managing CCA
component projects.

\csdms{} has developed innovative model/component interfaces that promote
model reuse and interoperability,
including the Basic Model Interface (\bmi{} - Section~\ref{sec:bmi}), which
uses
the \csdms{} Standard Names (Section~\ref{sec:standardnames}) and a framework
within which models can be coupled and enhanced with a set of \bmi{}-compatible
service components.
The \csdms{} Web Modeling Tool (\wmt{} - Section~\ref{sec:wmt}) is a web-based client that provides a
graphical interface to this framework and allows users to compose new models
by connecting and configuring components in a simple, browser-based graphical
interface.

\begin{comment}
In order to realize the full power and benefits of component-based modeling, a
modeling framework like \csdms{} needs an efficient mechanism to convert as many
open-source models as possible to reusable plug-and-play components. Since this
necessarily requires some involvement from each model's developer (\emph{who,
likely, is not paid to do this}), this mechanism must require minimal effort
and the benifit to the developer should be clear.
In the first few years of the \csdms{} project, we worked directly with model
developers and conducted one-on-one interviews to hear their concerns and
complaints about early designs and other frameworks. In addition to being easy
to implement, the following requirements of a model wrapping procedure became
clear:
%Through working directly with model developers and listening to their concerns
%and complaints about early designs and other frameworks,
%the following requirements presented themselves:
\begin{itemize}

%\item require minimal effort from the model developer,

\item allow the model to continue to be used in a stand-alone manner,

\item not introduce new dependencies into the model,

\item not interfere with the developer's design,

\item not require any modeling framework-specific knowledge on the part of the
      developer,

\item not require the addition of new code which accommodates the needs of
      other models.

\end{itemize}

%These requirements became clear during the first few years of the \csdms{}
%project
%by working directly with model developers and listening to their concerns and
%complaints about early designs and other frameworks.
\end{comment}

\section{Organizing and Documenting Open-Source Models}
\label{sec:organizing}

\subsection{The \csdms{} Model Repository}
\label{sec:repository}

Open-source software reduces redundancy, provides transparency through
external review and makes research replicable, which is fundamental to
scientific practices~\cite{ince2012case}.
Software accessibility is a key value of the \csdms{} mission. \csdms{}
ensures code developed by individual researchers or small research teams
remains accessible beyond the lifetime of a specific research project. Code
developers are required to distribute their codes under open-source licenses.

\csdms{} has built an online model repository that not only contains model
source code but also documents the submitted code as model metadata, runs
basic smoke-tests of the models, (optionally) maintains the source code in a
\csdms-hosted version control system, and provides easy downloads of the
original source. The repository now contains over 200 models and tools (over
4M lines of code) and has recieved over 13k model downloads.

\subsection{Digital Object Identifiers for Numerical Models}
\label{sec:doi}

\csdms{} has implemented a
Digital Object Identifier (\doi{}) system for software within the model
repository to guarantee recognition to software contributors.
\csdms{} was among the first software venues to assign \doi{}s to
open source software. The advantages of adopting a \doi{} system for open
source software are: 
\begin{itemize}
\item Guarantee credit to the software developer
\item Easily reuse and replicate research as software is directly locatable 
\item Increase software visibility, \doi{} content is 5 times more likely to
      deliver active links than content without.
\item Provide e.g. funding agencies the ability to track usage, so measure
      impact.
\end{itemize}

Only stable versions of open source
software that are physically hosted through the \csdms{} repository will be
assigned a \doi{}. For model software citations, \csdms{} recommends the
following the DataCite guidelines~\cite{brase2009datacite}.

\begin{shaded}
\leftskip 0.25in
\parindent -0.25in
\tt{
Developer, A., Developer, B. (Year of publication). \emph{Name of the model},
Model Version. Identifier.
}
\end{shaded}

\section{Reusable Components}
\label{sec:reusable}

\subsection{Sustainability and Interoperability Through Standards}
\label{sec:standards}

\subsubsection{The Basic Modeling Interface}
\label{sec:bmi}

The \emph{Basic Modeling Interface}\footnote{https://github.com/csdms/bmi}
(\bmi{}) is a component-coupling library
interface specification designed by \csdms~\cite{peckham2012component,
syvitski2014plug}.  In this context, a component is software that models a
particular environmental process and can be \emph{plugged into} another
component.

The \bmi{} specification was designed without any particular model-coupling
framework in mind.  Rather, it was designed to be framework agnostic. We
recognize model-coupling frameworks and tools may come and go but the
functionality a framework requires from its components will be significantly
more long-lived. Thus, a component should outlive any framework that it might
operate within.

The \bmi{} identifies the entry points into software components that provides a
calling appication with the necessary level of control to connect components.
\csdms, as well as other modeling frameworks such as
ESMF~\cite{hill2004architecture}, OpenMI~\cite{gregersen2007openmi}, and
OMS~\cite{david2002object}, identifies an interface that, at a minimum,
provides functionality to \emph{initialize}, \emph{update}, and
\emph{finalize} a component. However, for more complicated (and 2-way)
coupling components must implement the \bmi{} methods providing data access,
and model metadata.

\csdms{} has found that \bmi{} is an acceptable target for model developers.
The use of \bmi{} has also dramatically reduced the effort required by
\csdms{} staff to create and maintain components. So long as a model component
strictly exposes a \bmi{}, it can automatically be incorporated into the
\csdms{} model-coupling framework where it can then connect to compatible
components to form larger models and obtain additional functionality through
\bmi{} compatible tools (for example, NetCDF output, grid mapping).

\subsubsection{\csdms{} Standard Names}
\label{sec:standardnames}

The \csdms{} \emph{Standard Names}
\footnote{https://github.com/csdms/standard\_names}
are a common language for variable names exchanged between models. They play
an important role in the \bmi{} as they provide a mapping of a model's
internal variable names to a common language used by the \bmi{} getter and
setter functions.

Most models require input variables and produce output variables. For model
components to be reusable and interoperable (either within the \csdms{}
framework or any framework in general)
a set of components become a complete model when every component is able to
obtain the input variables it needs from another component. The \csdms{}
Standard Names were developed to provide a practical solution to this
semantic mediation problem ~\cite{peckham2012component, syvitski2014plug}.
The \csdms{} Standard Names provide a comprehensive set of naming rules and
patterns for creating unique labels for model variables that are not specific
to any particular modeling domain.

\subsection{Connecting and Running Models Through the Web}
\label{sec:wmt}

The \csdms{} Web Modeling Tool\footnote{https://csdms.colorado.edu/wmt}
(\wmt{}) is a web
application that provides an Ajax client-side (the \wmt{}
client\footnote{https://github.com/csdms/wmt-client}) graphical interface
(figure~\ref{fig:wmt_screenshot}) and a RESTful
server-side database and API (the \wmt{} server\footnote{https://github.com/csdms/wmt}) that allows users to build and run coupled Earth system models on
a high-performance computing cluster from a web browser. \wmt{} was designed
with four objectives:
\begin{itemize}

\item \emph{Accessibility}. As a web-based application, if you have access
      to the Internet, you have access to \wmt{}.

\item \emph{Integration}. Easily hyperlink from \wmt{} to resources on the 
      \csdms{} modeling portal (model documentation, labs, lectures,
      tutorials)
      %tutorials and movies - or to other resources on the Internet.

\item \emph{Portability}. \wmt{} has a native JavaScript interface, so it can
      be accessed on any modern web browser

%\item \emph{Maintenance}. Because modern browsers tend to adhere to web
%      standards, which lead to fewer cross-compatibility issues than
%      operating systems, only one version of WMT needs to be developed
%      and maintained.

\end{itemize}
With \wmt{}, users can: \emph{run} single component models, build (2-way)
\emph{coupled models}, view and \emph{edit} model parameters, \emph{share}
saved models with others in the community.

\begin{comment}
\begin{itemize}

\item select and run a single component model

\item build a coupled model from multiple components

\item view and edit the parameters of the model components;

%\item save models to a server, where they can be accessed on any computer
%      connected to the Internet;

\item share saved models with others in the community;

%\item run a model by connecting to a remote HPCC where the components are
%      installed.

\end{itemize}
\end{comment}

\begin{comment}
\subsection{Software sustainability through expanded functionality}

Another important goal of CSDMS is to enable the rapid development and application of linked dynamical models tailored to specific landscape-basin evolution problems and at specific temporal and spatial scales. Interdisciplinarity and coupling is a critical step towards advancing predictive modeling~\cite{voinov2010community}.

CSDMS develops a modeling environment containing a suite of integrated, ever-improving software modules, with expanded functionality beyond the original stand-alone codes.

The modeling coupling framework is designed to catalyze Earth-surface process research by empowering a broad community of scientists, students and applied modelers with computing tools, streamlining the process of idea generation and hypothesis testing through linked surface dynamics models, and enabling rapid creation and coupling of models tailored to specific settings, scientific problems, and time scales
\end{comment}

\section{Education and Evaluation of \csdms{} Tools}
\label{sec:education}

%\wmt{} provides the ability to run models with an easy to use graphical
%interface and intends to serve novice and more experienced modelers alike.
%The WMT beta version was released in May 2014  as a sequel to the CMT.

\csdms{} constantly evaluates the effectiveness of its tools and standards
through user feedback from clinics and tutorials and one-on-one interviews
with \csdms{} members.  Modeling courses at the University of Colorado,
clinics at the International Summer Institute at the University of Minnesota,
and clinics at annual CSDMS Annual Meetings are the dominant forums for
this testing and feedback.

Over the last 4 years over 150 graduate students - from geography, hydrology,
geology, oceanography, earth sciences and engineering departments - have
enrolled in intensive \csdms{}-run modeling courses. These courses
provide an excellent use case for us to test usability and access to the
tool. Students reported in post-course surveys that they were unfamiliar with
the use of models on a high performance computing cluster beforehand (rated
their familiarity as 1.5 average on a scale of 1-5), but ended the courses
with familiarity and being more comfortable with running and changing
simulations as well as visualizing results (average rating approaches 4 on
a Likert scale of 1-5)~\cite{overeem2013strategies}.

Course and clinic materials are posted online as educational resources.
\csdms{} develops \wmt{}-specific help pages for all models currently
available as components, so that users can easily learn about model
parameters and input files. In addition, we are developing web-based video
tutorials that show how to run models.

\section{Summary: Building an active and efficient modeling community}

\csdms{} provides cyber-infrastructure to promote the quantitative modeling of
earth surface processes. The US National Science Foundation funds \csdms{}, but
the community is international and includes industry and federal agency
representatives in addition to members from academic institutions. The
initiative has been growing with more than 150 people per year. \csdms{} now
fosters a community of approximately 1200 scientists who work on prediction
of the movement of fluids, and the production, erosion, transport, and
deposition of sediment and nutrients in landscapes and seascapes.

\section{Acknowledgements}

The \csdms{} Integration Facility operates under continuing grant 0621695
from the US National Science Foundation.

\begin{figure}
  \caption{The \csdms Web Modeling Tool}
  \begin{center}
    \includegraphics[scale=.25]{wmt.eps}
  \end{center}
  \label{fig:wmt_screenshot}
\end{figure}

\bibliographystyle{plain}

\bibliography{wssspe-citations}{}

\begin{comment}
\huttonnote{Add something about model DOIs in the abstract}

\huttonnote{
\begin{itemize}
\item improving the development process that leads to new software
  \begin{itemize}
  \item methods to develop sustainable software from the outset
  \item effective approaches to reusable software created as a by-product of
        research
  \item impact of computer science research on the development of scientific software
  \end{itemize}
\end{itemize}
}

\huttonnote{This is Irina's section.}

\huttonnote{
\begin{itemize}
\item building large and engaged user communities
  \begin{itemize}
  \item developing strong advocates
  \item measurement of usage and impact
  \end{itemize}
\item encouraging industry's role in sustainability
  \begin{itemize}
  \item engagement of industry with volunteer communities
  \item incentives for industry
  \item incentives for community to contribute to industry-driven projects
  \end{itemize}
\item improving education and training
  \begin{itemize}
  \item best practices for providing graduate students and postdoctoral
        researchers in domain communities with sufficient training in
        software development
  \item novel uses of sustainable software in education (K-20)
  \item case studies from students on issues around software development in
        the undergraduate or graduate curricula
  \end{itemize}
\end{itemize}
}

\huttonnote{This is Albert's section (he's rewriting this)}

\huttonnote{
\begin{itemize}
\item recommending policy changes
  \begin{itemize}
  \item software credit, attribution, incentive, and reward
  \item issues related to multiple organizations and multiple countries, such
        as intellectual property, licensing, etc.
  \item mechanisms and venues for publishing software, and the role of
        publishers
  \end{itemize}
\end{itemize}
}
\end{comment}

\end{document}